Bayesian Analysis of Gas-free Galaxies in Simulations I: Reviving Stochasticity

### 2022-10-23 papers/paper-1/sections/appendix.tex
Update figures to reflect hernquist runs
Update results section
Small updates to paper for distribution to Portegies-Zwart

\begin{figure*}
    \includegraphics[width=\textwidth]{{graham_density-A-C-3.0-0.05_prior_pred_log10_proj_density_mean}.png}
    \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit. Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs). The observed data is overlaid. From the prior predictive estimate, it is seen that the model and the assumed parameter hyperdistributions are able to describe the observed data, in that the observed data lies within the range allowed by the prior predictive estimate.}}
    \label{fig:priorpred}
\end{figure*}

\begin{figure*}
    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_0}.png}
    \caption{\drafting{Corner plot of hierarchical model hyperparameters. Each of the hyperdistributions are modelled as Gamma distributions with shape parameters $\alpha$ and $\beta$. For example, the hyperdistribution on the core radius $r_\mathrm{b}$ is modelled as $r_\mathrm{b} \sim \text{Gamma}(r_{\mathrm{b},\alpha}, r_{\mathrm{b},\beta})$. Critically, the hyperdistributions are unimodal with highest-density regions located some distance from the distribution edges, indicating the HMC sampling has converged well. }}
    \label{fig:corner_params_1}
\end{figure*}

\begin{figure*}
    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_1}.png}
    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
    \label{fig:corner_params_2}
\end{figure*}

\begin{figure*}
    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_2}.png}
    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
    \label{fig:corner_params_3}
\end{figure*}

>>>>

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_gqs_prior}.pdf}
    \caption{\drafting{Prior predictive distributions of latent parameters. Note the mode of the distributions roughly coincide with what one would expect from the literature.}}
    \label{fig:prior_latent}
\end{figure*}

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_prior_pred_log10_proj_density_mean}.pdf}
    \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit. Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs). The observed data is overlaid. From the prior predictive estimate, it is seen that the model and the assumed parameter hyperdistributions are able to describe the observed data, in that the observed data lies within the range allowed by the prior predictive estimate.}}
    \label{fig:priorpred}
\end{figure*}

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_pair_0}.pdf}
    \caption{\drafting{Corner plot of hierarchical model hyperparameters. Each of the hyperdistributions are modelled as Gamma distributions with shape parameters $\alpha$ and $\beta$. For example, the hyperdistribution on the core radius $r_\mathrm{b}$ is modelled as $r_\mathrm{b} \sim \text{Gamma}(r_{\mathrm{b},\alpha}, r_{\mathrm{b},\beta})$. Critically, the hyperdistributions are unimodal with highest-density regions located some distance from the distribution edges, indicating the HMC sampling has converged well. The contours correspond to the 50\%, 90\%, 95\%, and 99\% BCIs. }}
    \label{fig:corner_params_1}
\end{figure*}

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_pair_1}.pdf}
    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
    \label{fig:corner_params_2}
\end{figure*}

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_pair_2}.pdf}
    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
    \label{fig:corner_params_3}
\end{figure*}

### 2022-10-23 papers/paper-1/sections/intro.tex
Edits to paper 1
Small updates to paper for distribution to Portegies-Zwart

\drafting{Something cool}

\drafting{
    Introduce core sersic profile here.
}
\begin{equation}\label{eq:coresersic}
    \Sigma(R) = \Sigma'\left[ 1+\left(\frac{r_\mathrm{b}}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[ -b_n \left(\frac{R^\alpha+r_\mathrm{b}^\alpha}{R_\mathrm{e}^\alpha}\right)^{1/(\alpha n)} \right],
\end{equation}
where
\begin{equation}\label{eq:coresersic2}
    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[ b_n \left( 2^{1/\alpha} r_\mathrm{b}/R_\mathrm{e} \right)\right]
\end{equation}

>>>>

\drafting{There was a paper by \citet{nasim20} about ``defeating stochasticity''. They use some questionable methods to extrapolate some orders of magnitude to say that at physical resolution (i.e. solar mass particles) there is no stochasticity in binary evolution. I'm of the opinion this is not the case. The title of this paper is sort of a rebuke to their work.}

\drafting{
    Introduce necessary formulae
}
\begin{equation}\label{eq:coresersic}
    \Sigma(R) = \Sigma'\left[ 1+\left(\frac{r_\mathrm{b}}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[ -b_n \left(\frac{R^\alpha+r_\mathrm{b}^\alpha}{R_\mathrm{e}^\alpha}\right)^{1/(\alpha n)} \right],
\end{equation}
where
\begin{equation}\label{eq:coresersic2}
    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[ b_n \left( 2^{1/\alpha} r_\mathrm{b}/R_\mathrm{e} \right)\right]
\end{equation}

\begin{equation}\label{eq:quinlan_H}
    \dv{}{t}\left( \frac{1}{a} \right) = H \frac{G\rho}{\sigma}
\end{equation}

\begin{equation}\label{eq:quinlan_K}
    \dv{e}{t} = -K a^{-1} \dv{a}{t}
\end{equation}

\begin{equation}\label{eq:kepler_L}
    L = \mu\sqrt{GMa \left(1-e^2\right)}
\end{equation}

\begin{equation}\label{eq:influence_radius}
    r_\mathrm{infl} = r_\star(<2 M)
\end{equation}

\begin{equation}\label{eq:a_hard}
    a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{infl}}{4},
\end{equation}
where $q=M_{\bullet,2}/M_{\bullet,1} < 1$.

### 2022-10-23 papers/paper-1/sections/results.tex
Small updates to paper for distribution to Portegies-Zwart

\drafting{ENSURE CONSISTENCY WITH SIGMA AND I. AS WE ARE USING MASS DENSITY, SHOULD BE SIGMA, NOT I, OR ASSUME A M/L RATIO OF 1. NEED TO GO THROUGH AND ENSURE THIS}

@@

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_latentqty_compare}.pdf}
    \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats.}}
    \label{fig:cs_compare_naive}
\end{figure*}

@@

As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
We show an example of these latent parameter distributions in \autoref{fig:cs_compare_naive}, where the PDF of the latent parameter distributions is given in blue. 
For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive Gaussian statistics.
To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the mean and standard deviation of the sample. 
The mean value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
The Gaussian statistics fail to capture the correct estimates for the latent parameter estimates: this is seen most clearly for the latent parameter $n$, whose Gaussian error estimate implies that by $2\sigma$, there is a non-zero probability of $n$ assuming a value less than 0, which is unphysical.

We show the latent parameter distributions for each mass resolution family in \autoref{fig:cs_mass_compare}, and compare the median and IQRs \drafting{ensure this acronym explained somewhere} in \autoref{fig:cs_mass_compare_2}. 
We find first that the spread in the latent parameters (as quantified through their IQR) does not dramatically differ between mass resolutions.
Secondly, we observe that not all latent parameters in the core-S\"ersic model display asymptotic convergence, and that in some cases, such as for the effective radius $R_\mathrm{e}$, all tested mass resolutions produce estimates that are consistent with each other.
For other latent parameters, such as the core radius $r_\mathrm{b}$, there is a clear convergence of the median value with increasing mass resolution. \drafting{More particles --> core better resolved. Discuss here or in discussion?}

@@

>>>>

\drafting{ENSURE CONSISTENCY WITH BCI AND HDI. SAME THING, JUST DIFFERENT NAME}

@@

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_latentqty_compare}.pdf}
    \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats. Note error bars on black points are smaller than the point itself if not visible}}
    \label{fig:cs_compare_naive}
\end{figure*}

@@

As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
We show an example of these latent parameter distributions in \autoref{fig:cs_compare_naive}, where the PDF of the latent parameter distributions is given in blue. 
For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive Gaussian statistics.
To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the median and IQR of the sample. 
The median value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
The Gaussian statistics fail to capture the correct estimates for the latent parameter estimates, with some parameters (e.g. $n$) converging to the boundary value we limit the Least Squares method to (for $n$, the bounds are $[0,20]$).
We try successively relaxing and tightening the boundary values of each parameter to obtain a `physically-realistic' value for each optimal latent parameter value, with no success. 
Henceforth, we abandon any attempt to use a Least Squares minimisation to recover the optimal latent parameter values, returning instead to our Bayesian approach.

We show the latent parameter distributions for each mass resolution family in \autoref{fig:cs_mass_compare}, and compare the median and IQRs \drafting{ensure this acronym explained somewhere} in \autoref{fig:cs_mass_compare_2}. 
We find first that the spread in the latent parameters (as quantified through their IQR) does not dramatically differ between mass resolutions.
Secondly, we observe that not all latent parameters in the core-S\'ersic model display asymptotic convergence, and that in some cases, such as for the effective radius $R_\mathrm{e}$, all tested mass resolutions produce estimates that are consistent with each other.
For other latent parameters, such as the core radius $r_\mathrm{b}$, there is a clear convergence of the median value with increasing mass resolution. \drafting{More particles --> core better resolved. Discuss here or in discussion?}

@@

\drafting{
    QUESTIONS FOR SIMON:\\
    \begin{itemize}
        \item Hierarchical models rely on the ability to model a set of observations as belonging to a \textit{population}. After the chaos talk, can a collection of BH-perturbed systems be described still as belonging to a population? I think yes (stellar mass distribution unchanged), but would be keen to hear thoughts.
        \item In this paper we're perturbing the phase space coordinates of the BHs as opposed to a single field particle (we did this as another sensitivity test). Let's consider the case of perturbing a field particle. I find it still somehow curious that such a small displacement (a fraction of a softening length) still propagates the perturbation to such large scales, considering we are using a tree code for the particles not-near the BHs. I would have (na\"ively) thought that round off error from the tree algorithm would dominate the perturbation of a single particle (our code is deterministic given a specific hardware configuration, i.e. running the same ICs on the same CPU count will give identical results). Thoughts on this?
    \end{itemize}
}
